---
title: "d13C CSIA Analysis"
author: "Caila Kucheravy"
date: "`r Sys.Date()`"
output: pdf_document
---

Load packages and prep environment: 
```{r, echo = TRUE, message = FALSE, results = 'hide', warning = FALSE}
setwd("~/Documents/Master's/Analysis/CSIA")

library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stats)
library(lubridate)
library(ggfortify)
library(brms)
library(factoextra)
library(bayesplot)
library(tidybayes)
library(cowplot)
library(MASS)
```

Load d13C amino acid data & prep for analysis: 
```{r}
# Load data including means and se 
d13C_AA <- read.csv("input/d13C_AA.csv", header = TRUE) 

# Prep data
d13C_means <- d13C_AA %>% 
  # Filter for means
  filter(mean_se == "mean") %>% 
  # Create new column for "Group"
  unite(Group, c("Location", "Year"), sep = " ", remove = FALSE) %>% 
  # Remove AAs not considered: 
  dplyr::select(!all_of(c("His", "Lys", "Ser", "Tyr")))

# Add a column converting calendar date to julian day
d13C_means$Sample_date <- as.Date(d13C_means$Sample_date, "%Y-%m-%d")

d13C_means <- d13C_means %>% 
  mutate(julian_day = yday(d13C_means$Sample_date)) %>% 
  relocate(julian_day, .after = Sample_date)

# In the data there are same year and across-year duplicates - note that some samples have both within year and across year duplicates
# Same year duplicate coding: 
# 0 = no duplicate 
# 1 = duplicate sample, first collected 
# 2 = duplicate sample, second collected (excluded in analysis)

# Across year duplicate coding: 
# 0 = no duplicate 
# 1 = duplicate sample, first collected (earlier year)
# 2 = duplicate sample, second collected (sometimes same year as 1 if there is a 3, sometimes different year than 1)
# 3 = duplicate sample, third collected (sometimes same year as 2, sometimes different year than 1 & 2)

# Filter out same year and across year duplicates 
d13C_dups_removed <- d13C_means %>% 
  filter(!duplicate_in_year == "2") %>% 
  filter(!duplicate_other_year > 1)
```

## Mean standard deviation 

Calculate the mean standard deviation of replicate samples for methods: 
```{r}
d13C_se <- d13C_AA %>% 
  # Filter for means
  filter(mean_se == "se") %>% 
  # Remove samples from Sanikiluaq
  filter(!Location == "SANIKILUAQ") %>% 
  # select only AA columns
  dplyr::select(Ala, Asx, Glx, Gly, Ile, Leu, Lys, Met, Phe, Pro, Thr, Val)

mean_sd = mean(as.matrix(d13C_se[1:80,1:12]), na.rm = TRUE)
mean_sd
```

## Data Exploration

Boxplots for all d15N CSIA data.

Re-arrange data:
```{r}
# Essential AAs include Ile, Leu, Met, Phe, Val
# Sometimes threonine, but excluded here because of differences in beluga tissues

# AAs without all values: 
drop.cols.d13 <- c("Tissue", "UCDavis", "duplicate_in_year", "duplicate_other_year", "duplicate_id", "wgs", "sex")

# Rearrange the data for plotting 
d13N_rearrange <- d13C_dups_removed %>% 
  dplyr::select(!all_of(drop.cols.d13)) %>% 
  pivot_longer(11:21)

# Order AAs for plotting
d13N_rearrange$name <- factor(d13N_rearrange$name, levels = c("Ile", "Leu", "Met", "Phe", "Val", "Ala", "Asx", "Glx", "Gly", "Pro", "Thr"))
```

Boxplot by genetic group: 
```{r}
# separate by genetic group
outline_cols <- c("#0072B2", "#E69F00")
fill_cols <- c("#c6eaff", "#fff1d2")

# On one plot, shading essential amino acids 
ggplot(data = d13N_rearrange, aes(x = name, y = value, fill = genetic_group, colour = genetic_group)) + 
  geom_rect(aes(xmin = 0.5, xmax = 5.5, ymin = -Inf, ymax = Inf), fill = "grey93", colour = NA) +
  geom_boxplot() + 
  ylab(expression(paste(delta^{13}, "C", " (\u2030)"))) +
  scale_fill_manual(values = fill_cols, name = "Genetic Group") + 
  scale_colour_manual(values = outline_cols, name = "Genetic Group") +
  annotate(geom = "text", label = "Essential AAs", x = 3, y = -2, size = 5) +
  annotate(geom = "text", label = "Non-essential AAs", x = 8.5, y = -2, size = 5) +
  theme_classic() + 
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16))

#ggsave("plots/boxplots_d13C_allAAs.jpg", dpi = 300, width = 8, height = 5)
```

Essential amino acids: 
```{r}
remove <- c("Tissue", "UCDavis", "csia_sample_id", "Ala", "Asx", "Glx", "Gly", "Pro", "Thr")
 
d13C_AAess <- d13C_means %>% 
  dplyr::select(!all_of(remove))

# Filter out same year and across year duplicates 
d13C_AAess_dups_removed <- d13C_AAess %>% 
  filter(!duplicate_in_year == "2") %>% 
  filter(!duplicate_other_year > 1)
```

## Genetic and annual groups:

Run PCA with duplicates removed: 
```{r}
# PCA
d13C_AAess_pca <- prcomp(d13C_AAess_dups_removed[ ,15:19], scale = TRUE, center = TRUE)

# Scree plot: 
fviz_eig(d13C_AAess_pca)

# Biplot
biplot <- fviz_pca_biplot(d13C_AAess_pca, 
                repel = TRUE,
                col.var = "#FC4E07", # Variables color
                col.ind = "grey80",  # Individuals color
                label = "var",
                title = "")

biplot + 
  theme_classic() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12))

#ggsave("plots/d13C_AAess_pca.jpg", dpi = 300, width = 7, height = 5)
```

Plot to see genetic group:
```{r}
genetic_group_plot <- autoplot(d13C_AAess_pca, 
                               data = d13C_AAess_dups_removed,
                               colour = "genetic_group",
                               loadings = TRUE,
                               loadings.label = TRUE,
                               loadings.label.repel = TRUE,
                               frame = TRUE) 

genetic_group_plot + 
  scale_color_manual(values = outline_cols, name = "Genetic Group") + 
  scale_fill_manual(values = fill_cols, name = "Genetic Group") + 
  theme_classic() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12))

#ggsave("plots/pca_genetic_group_loadings.jpg", dpi = 300, width = 7, height = 5)
```

Take a look also at PC2 and PC3: 
```{r}
dimensions <- d13C_AAess_dups_removed %>% 
  dplyr::select(Group, DFO_sample_id, genetic_group) %>% 
  mutate(PC1 = d13C_AAess_pca$x[,1],
         PC2 = d13C_AAess_pca$x[,2],
         PC3 = d13C_AAess_pca$x[,3],
         PC4 = d13C_AAess_pca$x[,4])

ggplot(data = dimensions, aes(x = PC2, y = PC3, colour = genetic_group, fill = genetic_group)) + 
  geom_point() + 
  stat_ellipse() +
  scale_color_manual(values = outline_cols, name = "Genetic Group") + 
  scale_fill_manual(values = fill_cols, name = "Genetic Group") + 
  theme_classic()
```

Plot to see annual groups: 
```{r}
group_cols      <- c("#4100b3", "#004ebe", "#0071b3", "#008a9c", "#009e74", "#79a930", "#e6d700", "#e69d00", "#cf6f00", "#b34100", "#b30017", "#98005e")
group_fill_cols <- c("#dbc7ff", "#d2e4ff", "#c7eaff", "#d7faff", "#c5fff0", "#e6f2d2", "#fffcd2", "#fff1d2", "#ffe9cf", "#ffdbc7", "#ffc7ce", "#ffd3ee")

# Plot by group (location/year)
group_plot <- autoplot(d13C_AAess_pca, 
                       data = d13C_AAess_dups_removed,
                       colour = "Group",
                       frame = TRUE) 

group_plot + 
  scale_color_manual(values = group_cols) +
  scale_fill_manual(values = group_fill_cols) +
  theme_classic() + 
  #geom_text_repel(aes(label = DFO_sample_id), size = 1)+
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 12)
        )

#ggsave("plots/pca_annual_group.jpg", dpi = 600, width = 7.75, height = 5)
```

Since SPM and NFLD whales show a bit of a subgroup in the genetics data, we can see if they differ ecologically too.

Plot to see genetic sex:
```{r}
fill_cols_sex <- c("#ff8b8b", "#8bc5ff")
outline_cols_sex <- c("#8b0000", "#00468b")

genetic_group_plot <- autoplot(d13C_AAess_pca, 
                               data = d13C_AAess_dups_removed,
                               colour = "sex",
                               frame = TRUE) 

genetic_group_plot + 
  scale_color_manual(values = outline_cols_sex) + 
  scale_fill_manual(values = fill_cols_sex) + 
  theme_classic() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12))

#ggsave("plots/pca_sex.jpg", dpi = 300, width = 7, height = 5)
```

## Genetic subgroup - Newfoundland/St Pierre et Miquelon

The genetic PCA showed Newfoundland/St Pierre et Miquelon separating out from the Pond Inlet/Pangnirtung whales, suggesting genetic substructure. However, we only have a few SI samples from that area, so difficult to make any conclusions about diet/distribution from CSIA data. 

Separate out subgroups within the ECAG1 genetic group: 
```{r}
# Extract PCs from d13C AEss PCA: 
princomps <- d13C_AAess_pca$x

# Metadata: 
metadata_aess <- d13C_AAess_dups_removed %>% 
  dplyr::select(DFO_sample_id, Group, genetic_group, Location, Year)

# Add metadata: 
d13C_AEss_princomp1 <- cbind(metadata_aess, princomps)

# Identify individuals in subgroup
subgroups <- c("St. P & M 2021", "Newfoundland 2022")

# Add column for the diet clusters
d13C_AEss_princomps <- d13C_AEss_princomp1 %>% 
  mutate(subgroup = if_else(
    genetic_group == "ECAG2", "ECAG2", if_else(
      Group %in% subgroups, "ECAG1 Southern Subgroup", "ECAG1"
    )
  )) %>% 
  relocate(subgroup, .before = Location)

# Check if that worked 
subgroup_cols <-      c("#0072B2", "#009e73", "#E69F00")
subgroup_fill_cols <- c("#c6eaff", "#c5ffef", "#fff1d2")

ggplot(d13C_AEss_princomps, aes(x = PC1, y = PC2, color = subgroup)) + 
  geom_point() + 
  scale_color_manual(values = subgroup_cols) +
  theme_classic()

# Save RDS: 
#saveRDS(d13C_AEss_princomps, "output/d13C_AEss_subgroup_pcs.RDS")
```

```{r}
# Read in data if not running whole script
#d13C_AEss_princomps <- readRDS("output/d13C_AEss_subgroup_pcs.RDS")

# Extract only subgroups and sample ID: 
subgroups <- d13C_AEss_princomps %>%
  dplyr::select(DFO_sample_id, subgroup)
# saveRDS(subgroups, "output/genetic_subgroup_ids.RDS")
# write.csv(subgroups, "output/genetic_subgroup_ids.csv")
# subgroups <- readRDS("input/genetic_subgroup_ids.RDS")

# Add subgroup to dataset: 
d13C_AAess_dups_removed_subgroups <- d13C_AAess_dups_removed %>% 
  left_join(subgroups, by = "DFO_sample_id") %>% 
  relocate(subgroup, .after = "genetic_group")
```

Plot PCA with subgroups:
```{r}
subgroup_plot <- autoplot(d13C_AAess_pca, 
                          data = d13C_AAess_dups_removed_subgroups,
                          colour = "subgroup",
                          frame = TRUE) 

subgroup_plot + 
  scale_color_manual(values = subgroup_cols, name = "Subgroup") + 
  scale_fill_manual(values = subgroup_fill_cols, name = "Subgroup") + 
  theme_classic() + 
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12))

#ggsave("plots/pca_subgroup.jpg", dpi = 300, width = 7, height = 5)
```

Boxplots:
```{r}
d13N_AAess_rearrange <- d13C_AAess_dups_removed_subgroups %>% 
  pivot_longer(16:20)

ggplot(data = d13N_AAess_rearrange, aes(x = name, y = value, fill = subgroup, colour = subgroup)) + 
  geom_boxplot() + 
  ylab(expression(paste(delta^{13}, "C", " (\u2030)"))) +
  scale_fill_manual(values = subgroup_fill_cols, name = "Subgroup") + 
  scale_colour_manual(values = subgroup_cols, name = "Subgroup") +
  theme_classic() + 
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16))

#ggsave("plots/boxplot_group.jpg", dpi = 300, width = 8, height = 5)
```


## Inter-annual variation 

Some of these animals were re-sampled in different years. Let's run the PCA again with across-year duplicates to see how the values compare:
```{r}
# Colors for 7 duplicates: 
duplicate_cols <- c("#0071b3", "#696fc6", "#ac64c1", "#df58a5", "#fd5c7a", "#ff7748", "#e69d00")

# Label only certain points
library(ggrepel)

d13C_AAess_label_duplicates <- d13C_AAess %>% 
  filter(!duplicate_in_year == 2) %>% 
  mutate(label1 = if_else(duplicate_other_year > 0, DFO_sample_id, NA)) %>% 
  mutate(label2 = if_else(duplicate_other_year > 0, duplicate_id, NA))

# Run PCA with across year duplicate samples: 
d13C_AAess_pca_dups <- prcomp(d13C_AAess_label_duplicates[ ,15:19], scale = TRUE, center = TRUE)

# Plot
duplicate_pca <- autoplot(d13C_AAess_pca_dups, 
                          data = d13C_AAess_label_duplicates,
                          colour = "label2",
                          label = TRUE,
                          label.label = "label1",
                          label.repel = TRUE,
                          label.size = 3,
                          frame = FALSE) 

# Make nice
duplicate_pca + 
  geom_line(aes(group = duplicate_id, colour = label2)) +
  scale_color_manual(values = duplicate_cols) + 
  theme_classic() + 
  theme(legend.position='none',
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16)) 

#ggsave("plots/pca_duplicate.jpg", dpi = 300, width = 5.6, height = 5)
```

Some of the genetic duplicates are showing different grouping in different years. 

# Stats

## Linear discriminant analysis 

Standardize data: 
```{r}
# Scale data:
scaled_data <- as.data.frame(scale(d13C_AAess_dups_removed[,15:19]))

# Add sample IDs back to standardized data: 
scaled_data$DFO_sample_id <- d13C_AAess_dups_removed$DFO_sample_id

# Join scaled data with metadata sheet 
d13C_AAess_scaled <- d13C_AAess_dups_removed %>% 
  dplyr::select(Group, DFO_sample_id, genetic_group) %>% 
  left_join(scaled_data, by = "DFO_sample_id")
```

Discriminant analysis: 
```{r}
# Run discriminant analysis; genetic groups by the 5 AAess
d13C_AAess_lda <- lda(data = d13C_AAess_scaled,
                            genetic_group ~ Ile + Leu + Met + Phe + Val)

# See results 
d13C_AAess_lda
```

Extract fitted (predicted) values: 
```{r}
d13C_AAess_lda_pred <- predict(d13C_AAess_lda, d13C_AAess_scaled)
```

See if the LDA correctly categorized the data: 
```{r}
class <- d13C_AAess_lda_pred$class
table <- table(d13C_AAess_scaled$genetic_group, class)
table

classifications <- as.data.frame(cbind(d13C_AAess_scaled$DFO_sample_id, d13C_AAess_scaled$genetic_group, class))
```

GRNL-KW-2021-01 is the one that was misclassified. 

Plot: 
```{r}
plot(d13C_AAess_lda)

# Check calculations from the fitted model, and equations
library(ggplot2)
library(ggpubr)

outline_cols <- c("#0072B2", "#E69F00")
fill_cols <- c("#c6eaff", "#fff1d2")

# Using the predicted values from the model object
d13C_AAess_lda_pred_x <- data.frame(d13C_AAess_lda_pred$x)

d13C_AAess_lda_pred_x$genetic_group <- d13C_AAess_scaled$genetic_group

ggplot(d13C_AAess_lda_pred_x, 
                 aes(x=genetic_group, y=LD1, colour=genetic_group, fill = genetic_group)) + 
  geom_violin() +
  geom_jitter(size = 0.5, width = 0.3) + 
  scale_color_manual(values = outline_cols) + 
  scale_fill_manual(values = fill_cols) +
  theme_classic() + 
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.position = "none")

# ggsave("plots/LDA_boxplots.jpg", dpi = 300, width = 6, height = 5)

ggplot(d13C_AAess_lda_pred_x, 
                 aes(x = LD1, color = genetic_group, fill = genetic_group)) + 
  geom_histogram(position = "identity") +
  scale_color_manual(values = outline_cols) +
  scale_fill_manual(values = fill_cols) +
  theme_classic() + 
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        legend.position = c(0.9, 0.9))
```


## Bayesian models

Let's test whether the genetic groups differ in AAess using a multivariate model with brms (equivalent of MANOVA).
Vignette here: https://cran.r-project.org/web/packages/brms/vignettes/brms_multivariate.html
More help: https://solomonkurz.netlify.app/blog/2019-02-16-bayesian-correlations-let-s-talk-options/

Use mvbind to tell brms that all AAess variables should be response variables.
Use set_rescor(TRUE) to fit model with residual correlations.

### Models

Model 1 - genetic_group:
```{r}
# Define the model
genetic_group_model <- bf(mvbind(Ile, Leu, Met, Phe, Val) ~ genetic_group) + set_rescor(TRUE)

# # Fit the model (on server)
# d13C_Aess__genetic_group <- brm(genetic_group_model,
#                                 data = d13C_AAess_dups_removed,
#                                 warmup = 20000,
#                                 iter = 50000,
#                                 chains = 3,
#                                 init = "random",
#                                 cores = 2)
# 
# # Save the model (on server)
# saveRDS(d13C_Aess__genetic_group, "output/d13C_Aess__genetic_group.rds")

# Load model and view results: 
d13C_Aess__genetic_group <- readRDS("server/d13C_Aess__genetic_group.rds")
summary(d13C_Aess__genetic_group)
```


Model 2 - genetic group + year:
```{r}
# Define the model
genetic_group_year_model <- bf(mvbind(Ile, Leu, Met, Phe, Val) ~ genetic_group + Year) + set_rescor(TRUE)

# # Fit the model (on server)
# d13C_Aess__genetic_group_year <- brm(genetic_group_year_model,
#                                      data = d13C_AAess_dups_removed,
#                                      warmup = 20000,
#                                      iter = 50000,
#                                      chains = 3,
#                                      init = "random",
#                                      cores = 2)
# 
# # Save the model (on server)
# saveRDS(d13C_Aess__genetic_group_year, "output/d13C_Aess__genetic_group_year.rds")

# Load model and view results: 
# d13C_Aess__genetic_group_year <- readRDS("server/d13C_Aess__genetic_group_year.rds")
# summary(d13C_Aess__genetic_group_year)
```

Here year is significant for every AA. But I think mostly influenced by the 2020 subgroup. 

Model 3.1 - subgroup:
```{r}
# Define the model
subgroup_model <- bf(mvbind(Ile, Leu, Met, Phe, Val) ~ subgroup) + set_rescor(TRUE)

# # Fit the model (on server)
# d13C_Aess__subgroup <- brm(subgroup_model,
#                            data = d13C_AAess_dups_removed_subgroups,
#                            warmup = 20000,
#                            iter = 50000,
#                            chains = 3,
#                            init = "random",
#                            cores = 2)
# 
# # Save the model (on server)
# saveRDS(d13C_Aess__subgroup, "output/d13C_Aess__subgroup.rds")

# Load model and view results: 
d13C_Aess__subgroup <- readRDS("server/d13C_Aess__subgroup.rds")
summary(d13C_Aess__subgroup)
```


Model 3.2 - subgroup + year:
```{r}
# Define the model
subgroup_year_model <- bf(mvbind(Ile, Leu, Met, Phe, Val) ~ subgroup + Year) + set_rescor(TRUE)

# # Fit the model (on server)
# d13C_Aess__subgroup_year <- brm(subgroup_year_model,
#                                 data = d13C_AAess_dups_removed_subgroups,
#                                 warmup = 20000,
#                                 iter = 50000,
#                                 chains = 3,
#                                 init = "random",
#                                 cores = 2)
# 
# # Save the model (on server)
# saveRDS(d13C_Aess__subgroup_year, "output/d13C_Aess__subgroup_year.rds")

# Load model and view results: 
# d13C_Aess__subgroup_year <- readRDS("server/d13C_Aess__subgroup_year.rds")
# summary(d13C_Aess__subgroup_year)
```

Model 4 - genetic group + Julian day:
```{r}
# Define the model
julian_day_model <- bf(mvbind(Ile, Leu, Met, Phe, Val) ~ genetic_group + julian_day) + set_rescor(TRUE)

# # Fit the model (on server)
# d13C_Aess__julian_day <- brm(julian_day_model,
#                              data = d13C_AAess_dups_removed,
#                              warmup = 20000,
#                              iter = 50000,
#                              chains = 3,
#                              init = "random",
#                              cores = 2)
# 
# # Save the model (on server)
# saveRDS(d13C_Aess__julian_day, "output/d13C_Aess__julian_day.rds")

# Load model and view results: 
d13C_Aess__julian_day <- readRDS("server/d13C_Aess__julian_day.rds")
summary(d13C_Aess__julian_day)
```

No relationship with Julian Day. 

Model 5 - location/year group within ECAG1:
```{r}
# Remove the ECAG2 group, east greenland, and st pierre 2021, and pang 2021 (since only one left after removing duplicates)
group_data <- d13C_AAess_dups_removed %>% 
  filter(!genetic_group == "ECAG2") %>% 
  filter(!Group %in% c("St. P & M 2021", "Newfoundland 2022", "Ittoqqortoormiit 2021"))

# Define the model
group_model <- bf(mvbind(Ile, Leu, Met, Phe, Val) ~ Group) + set_rescor(TRUE)

# # Fit the model (on server)
# d13C_Aess__group <- brm(group_model,
#                         data = group_data,
#                         warmup = 20000,
#                         iter = 50000,
#                         chains = 3,
#                         init = "random",
#                         cores = 2)
# 
# # Save the model (on server)
# saveRDS(d13C_Aess__group, "output/d13C_Aess__group.rds")

# Load model and view results: 
d13C_Aess__group <- readRDS("server/d13C_Aess__group.rds")
summary(d13C_Aess__group)
```


Model 6 - genetic group + sex:
```{r}
# Define the model
sex_model <- bf(mvbind(Ile, Leu, Met, Phe, Val) ~ genetic_group + sex) + set_rescor(TRUE)

# # Fit the model (on server)
# d13C_Aess__sex <- brm(sex_model,
#                       data = d13C_AAess_dups_removed,
#                       warmup = 20000,
#                       iter = 50000,
#                       chains = 3,
#                       init = "random",
#                       cores = 2)
# 
# # Save the model (on server)
# saveRDS(d13C_Aess__sex, "output/d13C_Aess__sex.rds")

# Load model and view results: 
d13C_Aess__sex <- readRDS("server/d13C_Aess__sex.rds")
summary(d13C_Aess__sex)
```

None of the AAs differ by sex. 

### Diagnostics & Posterior Predictive Checks

Model to be checked: 
```{r}
# model to evaluate 
model <- d13C_Aess__genetic_group   
# model name
model_name <- "d13C_Aess__genetic_group"
# Glx-Phe or Thr-Phe
values <- c("Ile", "Leu", "Met", "Phe", "Val")               
# genetic_group or subgroup
group <- "genetic_group"            
```


Check Gelman-Rubin Diagnostics: 
```{r}
gr_diag <- brms::rhat(model)
write.csv(gr_diag, file = paste("output/gelmanrubin_", model_name, ".csv", sep=""))
```


Traceplot and posterior density plots: 
```{r}
# Set colour palette for bayesplot
color_scheme_set("blue")

# Set parameters:
pars_intercept    = c("b_Ile_Intercept", "b_Leu_Intercept", "b_Met_Intercept", "b_Phe_Intercept", "b_Val_Intercept")
pars_geneticgroup = c("b_Ile_genetic_groupECAG2", "b_Leu_genetic_groupECAG2", "b_Met_genetic_groupECAG2", "b_Phe_genetic_groupECAG2", "b_Val_genetic_groupECAG2")
#pars_subgroup1    = c("b_Ile_subgroupECAG1_2", "b_Leu_subgroupECAG1_2", "b_Met_subgroupECAG1_2", "b_Phe_subgroupECAG1_2", "b_Val_subgroupECAG1_2")
#pars_subgroup2    = c("b_Ile_subgroupECAG2", "b_Leu_subgroupECAG2", "b_Met_subgroupECAG2", "b_Phe_subgroupECAG2", "b_Val_subgroupECAG2")
pars_sigma        = c("sigma_Ile", "sigma_Leu", "sigma_Met", "sigma_Phe", "sigma_Val")

# Traceplots:
# jpeg(filename = paste("plots/", model_name, "__traceplot_intercept.jpeg"), units="in", width = 8, height = 5, res = 400)
mcmc_trace(model, pars = pars_intercept, facet_args = list(ncol = 1, strip.position = "left"))
# dev.off()

# jpeg(filename = paste("plots/", model_name, "__traceplot_geneticgroup.jpeg"), units="in", width = 8, height = 5, res = 400)
mcmc_trace(model, pars = pars_geneticgroup, facet_args = list(ncol = 1, strip.position = "left"))
# dev.off()

# jpeg(filename = paste("plots/", model_name, "__traceplot_sigma.jpeg"), units="in", width = 8, height = 5, res = 400)
mcmc_trace(model, pars = pars_sigma, facet_args = list(ncol = 1, strip.position = "left"))
# dev.off()

# Density plots: 
# jpeg(filename = paste("plots/", model_name, "__densplot_intercept.jpeg"), units="in", width = 8, height = 2, res = 400)
mcmc_dens_overlay(model, pars = pars_intercept, facet_args = list(nrow = 1))
# dev.off()

# jpeg(filename = paste("plots/", model_name, "__densplot_geneticgroup.jpeg"), units="in", width = 8, height = 2, res = 400)
mcmc_dens_overlay(model, pars = pars_geneticgroup, facet_args = list(nrow = 1))
# dev.off()

# jpeg(filename = paste("plots/", model_name, "__densplot_sigma.jpeg"), units="in", width = 8, height = 2, res = 400)
mcmc_dens_overlay(model, pars = pars_sigma, facet_args = list(nrow = 1))
# dev.off()
```

Plot coefficients: 
```{r}
effects <- data.frame(fixef(model))

ggplot(effects) + 
  geom_vline(xintercept = 0, lty = 2, col = "#0071b3") +
  geom_point(aes(x = Estimate, y = rownames(effects))) + 
  geom_segment(aes(x = Q2.5, xend = Q97.5, y = rownames(effects), yend = rownames(effects))) + 
  theme_classic() + 
  theme(axis.title.y = element_blank())

#ggsave(paste("plots/", model_name, "__coefplot.jpeg"), dpi = 300, width = 8, height = 5)

effects_intercept <- effects %>% 
  filter(grepl("Intercept", rownames(effects)))

ggplot(effects_intercept) + 
  geom_vline(xintercept = 0, lty = 2, col = "darkred") +
  geom_point(aes(x = Estimate, y = rownames(effects_intercept))) + 
  geom_segment(aes(x = Q2.5, xend = Q97.5, y = rownames(effects_intercept), yend = rownames(effects_intercept))) + 
  theme_classic() + 
  theme(axis.title.y = element_blank())

effects_other <- effects %>% 
  filter(!grepl("Intercept", rownames(effects)))

ggplot(effects_other) + 
  geom_vline(xintercept = 0, lty = 2, col = "darkred") +
  geom_point(aes(x = Estimate, y = rownames(effects_other))) + 
  geom_segment(aes(x = Q2.5, xend = Q97.5, y = rownames(effects_other), yend = rownames(effects_other))) + 
  theme_classic() + 
  theme(axis.title.y = element_blank())
```

Conditional effects plot: 
```{r}
conditional_effects(model)
```

Autocorrelation plot: 
```{r}
mcmc_acf(model, pars = pars_geneticgroup)
#ggsave(paste("plots/", model_name, "__autocorrplot.jpeg"), dpi = 300, width = 8, height = 5)
```

Check model fit with posterior predictive check: 

PP check 1, density overlap plot:
```{r}
AAs <- c("Ile", "Leu", "Met", "Phe", "Val")

for (AA in AAs) {
  pp_check <- pp_check(model, 
                       type = "dens_overlay_grouped",
                       ndraws = 100, 
                       group = group,
                       resp = AA)
  # show plot
  print(pp_check)
  # save plot
  #ggsave(paste("plots/", model_name, "_", AA, "__ppcheck_densoverlay.jpeg", sep = ""), dpi = 300, width = 8, height = 5)
}
```

PP check 2, mean grouped: 
```{r, warning=FALSE}
for (AA in AAs) {
  pp_check2 <- pp_check(model, 
         type = "stat_grouped", 
         stat = "mean",
         ndraws = 100, 
         group = group,
         resp = AA)
  # Show plot:
  print(pp_check2)
  # Save plot:
  #ggsave(paste("plots/", model_name, "_", AA, "__ppcheck_meangrouped.jpeg", sep = ""), dpi = 300, width = 8, height = 5)
}
```

PP check 3 - stat 2d with mean and se:
```{r, warning=FALSE}
for (AA in AAs) {
  pp_check3 <- pp_check(model, 
         type = "stat_2d", 
         stat = c("mean", "sd"),
         resp = AA)
  # Show plot:
  print(pp_check3)
  # Save plot:
  #ggsave(paste("plots/", model_name, "_", AA, "__ppcheck_stat2d.jpeg", sep = ""), dpi = 300, width = 8, height = 5)
}
```

Visualize the d13C values per group with associated uncertainty:
```{r}
# Couldn't figure out how to do this one in a for loop, so have to enter the AAs manually

# AAs <- c("Ile", "Leu", "Met", "Phe", "Val")

plot_spreaddraws <- model %>%
  # spread 1000 posterior samples for 2 parameters in wide format
  spread_draws(b_Phe_Intercept, b_Phe_genetic_groupECAG2, ndraws = 1000, seed = 123) %>%
  # calculate average numbers and convert to long format for visualisation
  mutate(ECAG1 = b_Phe_Intercept,
         ECAG2 = b_Phe_Intercept + b_Phe_genetic_groupECAG2) %>%
  pivot_longer(cols = c("ECAG1", "ECAG2"), 
               names_to = "genetic_group",                
               values_to = "Phe") %>%
  # visualise via ggplot()
  ggplot(aes(y = Phe, x = genetic_group, color = genetic_group, fill = genetic_group)) +
    stat_eye(point_interval = "median_qi", .width = c(0.6, 0.9)) + 
    scale_color_manual(values = outline_cols) +
    scale_fill_manual(values = fill_cols) +
    theme_classic() + 
    theme(legend.position = "none", 
          axis.title.x = element_blank())
  
# Show plot:
print(plot_spreaddraws)
# Save plot:
#ggsave(paste("plots/", model_name, "_Phe__spreaddraws.jpeg", sep = ""), dpi = 300, width = 7, height = 5)
```

Hypothesis test of whether the two groups are different: 
```{r}
hypothesis <- hypothesis(d13C_Aess__genetic_group, "Ile_genetic_groupECAG2 = 0", alpha = 0.05)
hypothesis
plot(hypothesis)
```





